\setcounter{figure}{0}
\setcounter{listing}{0}

\chapter{Work schedule throughout semesters \label{cha:work_schedule} }

\section*{Winter semester}

\begin{table}[htpb]
    \centering
    \begin{tabular}{|c|l|}
    \hline
    \cellcolor[gray]{0.8}\textbf{Semester week number} & \cellcolor[gray]{0.8}\textbf{Description} \\ \hline

    \textbf{1} & \textit{Studying the problem} \\ \hline
    \textbf{2} & \textit{Finding literature} \\ \hline
    \textbf{3} & \textit{Finding more sources} \\ \hline
    \textbf{4} & \textit{EU AI Act analysis} \\ \hline
    \textbf{5} & \textit{Jailbreak analysis} \\ \hline
    \textbf{6} & \textit{AI analysis} \\ \hline
    \textbf{7} & \textit{Content filter analysis} \\ \hline
    \textbf{8} & \textit{Risks of AI solutions analysis} \\ \hline
    \textbf{9} & \textit{Experimenting with Jailbreaking} \\ \hline
    \textbf{10} & \textit{More experimenting with Jailbreaking} \\ \hline
    \textbf{11} & \textit{Evaluating of the experiments} \\ \hline
    \textbf{12} & \textit{Document revision and final changes} \\ \hline
    \end{tabular}
\end{table}

\subsection*{Plan evaluation}

The first part of my Bachelor's thesis (BP1) was done in the winter semester and in the following examination period. In the first weeks I was finding sources for my thesis, studying them, and finally started writing. Unfortunately, I set my expectations a little higher. The winter semester was tough and I needed to meet deadlines for other courses, so I was writing less than I intended to. Sometimes I did my analyzes of the mentioned problems in another order because I was more interested in them. My supervisor and I agreed to have consultations every week. I tried to be present every week, but sometimes I did not have time due to the busy semester. In each consultation, I have discussed progress with my supervisor, and he gave me valuable feedback on changes or additions that I could implement. Finally, I finished the first part of the thesis during the examination period, where I had time because my examinations were not until January. Some of the mentioned topics in the plan are set as TBD, and other topics which were not in this semester plan as well where I will be working on them in the summer semester as part of BP2 and the final thesis.


\newpage


\section*{Summer semester}

\begin{table}[htpb]
    \centering
    \begin{tabular}{|c|l|}
    \hline
    \cellcolor[gray]{0.8}\textbf{Semester week number} & \cellcolor[gray]{0.8}\textbf{Description} \\ \hline

    \textbf{1} & \textit{Revision and proofreading of the document} \\ \hline
    \textbf{2} & \textit{U.S. Legislation Analysis} \\ \hline
    \textbf{3} & \textit{China Legislation Analysis} \\ \hline
    \textbf{4} & \textit{Solution proposal} \\ \hline
    \textbf{5} & \textit{Experimenting with Jailbreaking} \\ \hline
    \textbf{6} & \textit{Experimenting with Jailbreaking} \\ \hline
    \textbf{7} & \textit{Experimenting with Jailbreaking} \\ \hline
    \textbf{8} & \textit{Evaluating of the experiments} \\ \hline
    \textbf{9} & \textit{Guidelines for users} \\ \hline
    \textbf{10} & \textit{Guidelines for users} \\ \hline
    \textbf{11} & \textit{Conclusion and Resumé} \\ \hline
    \textbf{12} & \textit{Document revision and final changes} \\ \hline
    \end{tabular}
\end{table}

\subsection*{Plan Evaluation}
The second part of my Bachelor's thesis (BP2) was completed during the summer semester. I started by reviewing and proofreading the document that I had created in the winter. I was able to fix some unclear wording, formatting, etc. Then I continued where I left off, which was the legislation. Then I wrote a solution proposal for the practical part of the thesis which were the guidelines. After that, I came up with several experiments that needed more time for execution. Following the experiments, I evaluated them and started working on the guidelines. Lastly, I wrote the conclusion for my thesis and translated the most important parts to Slovak for the Resumé as a requirement of writing the thesis in English. Finally, I reviewed the entire document, made final changes, and submitted it.
I was able to follow the plan with some slowdowns that were caused by my other assignments, and I was roughly a week and half behind the schedule that I needed to catch up with.


\begin{refsegment}

% Bibliography
\printbibliography[heading=referencessec,segment=\therefsegment,resetnumbers=true]

\end{refsegment}